---
title: "HMO_Data_Cleaning"
author: "Harshit"
date: "2022-11-14"
output: html_document
---


```{r}
#install.packages("tidyverse")
library(tidyverse)
file <- ("https://intro-datascience.s3.us-east-2.amazonaws.com/HMO_data.csv")
HMO_data <- read_csv(file)
#View(HMO_data)
```


```{r}
#DATA CLEANING
sum(is.na(HMO_data))
dim(HMO_data)
summary(HMO_data)
```


```{r}
# IMPUTATION
#install.packages("imputeTS")
library(imputeTS)
HMO_data$bmi <- na.interpolation(HMO_data$bmi)
#View(HMO_data)
sum(is.na(HMO_data$bmi))

HMO_data$hypertension <- na.interpolation(HMO_data$hypertension)
#View(HMO_data)
sum(is.na(HMO_data$hypertension))
```
```{r}
#HISTOGRAMS AND BAR PLOTS

#UNIVARIATE ANALYSIS
hist(HMO_data$age,labels = paste0(round(hist(HMO_data$age, plot = FALSE)$counts / length(HMO_data$age) * 100, 1), "%"))
```
```{r}
#the dataset has more proportion of people in their 20s (13.2%) followed by people in their late 40s to late 50s (10.8 and 10.7% respectively)
# and their are very few people above 65
```


```{r}
table1<- table(HMO_data$gender)
gender_prop<- prop.table(table1)
gender_prop<- data.frame(gender_prop)
gender_prop
```
```{r}
#the proportion of male is slightly more than females in the given sample```
```

```{r}
hist(HMO_data$cost,labels = paste0(round(hist(HMO_data$cost, plot = FALSE)$counts / length(HMO_data$cost) * 100, 1), "%"))
```
```{r}
#the cost for most of the people lie below 10000 (84.6%). Very few proportion of people have cost above 10000
```


```{r}
library(scales)
HMO_data %>% 
  count(gender) %>% 
  mutate(perc = n / nrow(HMO_data)) -> tips2

tips2

ggplot(tips2, aes(x = gender, y = perc,fill=gender)) + geom_bar(stat = "identity") +  geom_text(aes( label = scales::percent(..prop..),
                   y= ..prop.. ), stat= "count", vjust = -.5)
```
```{r}
#BIVARIATE ANALYSIS
plot_1 <- ggplot(data=HMO_data) + aes(x=age,y=cost) + geom_point() + geom_smooth(method="lm", se=FALSE)

plot_1 <- plot_1 + ggtitle("Scatterplot of cost vs age")

plot_1
```
```{r}
#the cost of people is increasing slightly with increase in age
```

```{r}
ggplot(HMO_data, aes(x = HMO_data$bmi)) + geom_histogram(aes(fill =HMO_data$location))
ggplot(HMO_data, aes(x = HMO_data$bmi)) + geom_histogram(aes(fill =HMO_data$location_type))
ggplot(HMO_data, aes(x = HMO_data$cost)) + geom_histogram(aes(fill =HMO_data$location))
ggplot(HMO_data, aes(x = HMO_data$cost)) + geom_histogram(aes(fill =HMO_data$location_type))
ggplot(HMO_data, aes(x = HMO_data$cost)) + geom_histogram(aes(fill =HMO_data$smoker))
ggplot(HMO_data, aes(x = HMO_data$bmi)) + geom_histogram(aes(fill =HMO_data$smoker))
ggplot(HMO_data, aes(x = HMO_data$cost)) + geom_histogram(aes(fill =HMO_data$education_level))
ggplot(HMO_data, aes(x = HMO_data$bmi)) + geom_histogram(aes(fill =HMO_data$education_level))

```


```{r}
ggplot(HMO_data, aes(x = HMO_data$children)) + geom_bar(aes(fill =HMO_data$gender),position = "dodge")
ggplot(HMO_data, aes(x = HMO_data$exercise)) + geom_bar(aes(fill =HMO_data$gender),position = "dodge")
ggplot(HMO_data, aes(x = HMO_data$smoker)) + geom_bar(aes(fill =HMO_data$gender),position = "dodge")
ggplot(HMO_data, aes(x = HMO_data$education_level)) + geom_bar(aes(fill =HMO_data$gender),position = "dodge")
ggplot(HMO_data, aes(x = HMO_data$location_type)) + geom_bar(aes(fill =HMO_data$gender),position = "dodge")
ggplot(HMO_data, aes(x = HMO_data$smoker)) + geom_bar(aes(fill =HMO_data$gender),position = "dodge")
```
```{r}

ggplot(HMO_data, aes(x = HMO_data$smoker, y=HMO_data$cost)) + geom_histogram(aes(fill =HMO_data$gender))
barplot(cost ~ gender,data=HMO_data)
boxplot(cost ~ smoker,data=HMO_data)
boxplot(cost ~ location,data=HMO_data)
boxplot(cost ~ location_type,data=HMO_data)
boxplot(cost ~ education_level,data=HMO_data)
```


```{r}
#CATEGORICAL EDA- TABLES
cost_exercise<- data.frame(HMO_data$exercise,HMO_data$cost,HMO_data$location)
active_cost<- cost_exercise[which(cost_exercise$HMO_data.exercise=='Active'),]
active_cost
aggregate(HMO_data.cost~ HMO_data.location, active_cost, mean)
```
```{r}
#among people who are exercising, MASSACHUSETTS has the most costly patients while CONNECTICUT has less costly patients
```

```{r}
notactive_cost<- cost_exercise[which(cost_exercise$HMO_data.exercise=='Not-Active'),]
notactive_cost
aggregate(HMO_data.cost~ HMO_data.location, notactive_cost, mean)
```
```{r}
#among people who are not exercising, NEW YORK has the most costly patients while MARYLAND has less costly patients
```

```{r}
cost <- ggplot(HMO_data, aes(y=cost)) + 
   geom_boxplot(outlier.colour="red", outlier.shape=8,
                outlier.size=4)
cost

```


```{r}
boxplot(cost ~ gender,data=HMO_data)
boxplot(cost ~ smoker,data=HMO_data)
boxplot(cost ~ location,data=HMO_data)
boxplot(cost ~ location_type,data=HMO_data)
boxplot(cost ~ education_level,data=HMO_data)

```


```{r}
boxplot(bmi ~ gender,data=HMO_data)
boxplot(bmi ~ smoker,data=HMO_data)
boxplot(bmi ~ location,data=HMO_data)
boxplot(bmi ~ location_type,data=HMO_data)
boxplot(bmi ~ education_level,data=HMO_data)
```

```{r}
cost_bmi<- data.frame(HMO_data$bmi,HMO_data$cost,HMO_data$location)
max_bmi_cost<- cost_bmi[which.max(cost_bmi$HMO_data.bmi),]
min_bmi_cost<- cost_bmi[which.min(cost_bmi$HMO_data.bmi),]
max_bmi_cost
min_bmi_cost 

#then max bmi is 53.13 with the cost of 331 in pennsylvania
#then min bmi is 15.96 with the cost of 213 in pennsylvania
```


```{r}
library(dplyr)

location_avg_cost<- HMO_data %>% 
  group_by(HMO_data$location) %>% 
  summarize(mean_age = mean(age), mean_cost = mean(cost))

location_avg_cost<-data.frame(location_avg_cost)
location_avg_cost<- location_avg_cost[order(location_avg_cost$mean_cost),]
location_avg_cost

```

```{r}
#the mean age of pateints in the sample states is around 38, with MASSACHUSETTS having the oldest mean age

#new york is having the highest mean cost, while CONNECTICUT has the least mean cost
```

```{r}
cost_smoker<- data.frame(HMO_data$smoker,HMO_data$cost,HMO_data$location)
smoker_cost1<- cost_smoker[which(cost_smoker$HMO_data.smoker=='yes'),]
smoker_cost1
aggregate(HMO_data.cost~ HMO_data.location, smoker_cost1, mean)

```
```{r}
# among people who smoke, the state of NEW YORK has the most cost, while MARYLAND has less cost
```

```{r}
cost_smoker<- data.frame(HMO_data$smoker,HMO_data$cost,HMO_data$location)
smoker_cost2<- cost_smoker[which(cost_smoker$HMO_data.smoker=='no'),]
smoker_cost2
aggregate(HMO_data.cost~ HMO_data.location, smoker_cost2, mean)
```
```{r}
# among people who do not smoke, the state of NEW YORK has the most cost, while CONNECTICUT has less cost
```


```{r}
#MAPS

library(ggplot2); 
library(maps); 
library(ggmap); 
library(mapproj)
us <- map_data("state") # fetches the state data column from the dataframe

location_avg_cost$HMO_data.location <- tolower(location_avg_cost$HMO_data.location)
MyNewPlot <- merge(location_avg_cost,us,by.x="HMO_data.location",by.y="region")
MyNewPlot <- MyNewPlot %>% arrange(order)

c <- ggplot(MyNewPlot)
c <- c + aes(x=long,y=lat,group = group) + geom_polygon(aes(fill= mean_cost), color = "black")
c <- c + expand_limits(x=MyNewPlot$long,y=MyNewPlot$lat) + coord_map(projection = "mercator")
c <- c+ scale_fill_gradient(low='white', high='blue')
c
```


```{r}
#table(list(HMO_data$location))
c2 <- ggplot(MyNewPlot)
c2 <- c2 + aes(x=long,y=lat,group = group) + geom_polygon(aes(fill= mean_age), color = "black")
c2 <- c2 + expand_limits(x=MyNewPlot$long,y=MyNewPlot$lat) + coord_map(projection = "mercator")
c2 <- c2+ scale_fill_gradient(low='white', high='blue')
c2

#the map shows the distribution of age in the seven states of the sample. 

```




```{r}
#ADDING EXPENSIVE COLUMN

HMO_data2 <- HMO_data
HMO_data2$expensive <- as.factor(HMO_data$cost>4775)
HMO_data2

#0-> not expensive
#1 -> expensive

```

```{r}
#EDA with expensive column

counts<- table(HMO_data2$expensive)

barplot(counts, main="ExpensiveDistribution",
   xlab="Expensive")

```
```{r}

ggplot(HMO_data2, aes(x = HMO_data2$children)) + geom_bar(aes(fill =HMO_data2$expensive),position = "dodge")

ggplot(HMO_data2, aes(x = HMO_data2$education_level)) + geom_bar(aes(fill =HMO_data2$expensive),position = "dodge")

ggplot(HMO_data2, aes(x = HMO_data2$smoker)) + geom_bar(aes(fill =HMO_data2$expensive),position = "dodge")

ggplot(HMO_data2, aes(x = HMO_data2$location)) + geom_bar(aes(fill =HMO_data2$expensive),position = "dodge")

ggplot(HMO_data2, aes(x = HMO_data2$exercise)) + geom_bar(aes(fill =HMO_data2$expensive),position = "dodge")

ggplot(HMO_data2, aes(x = HMO_data2$gender)) + geom_bar(aes(fill =HMO_data2$expensive),position = "dodge")

ggplot(HMO_data2, aes(x = HMO_data2$married)) + geom_bar(aes(fill =HMO_data2$expensive),position = "dodge")


```
```{r}
ggplot(HMO_data2, aes(x = HMO_data2$bmi)) + geom_histogram(aes(fill =HMO_data2$expensive))

ggplot(HMO_data2, aes(x = HMO_data2$age)) + geom_histogram(aes(fill =HMO_data2$expensive))


```
```{r}
boxplot(age ~ expensive,data=HMO_data2)
boxplot(bmi ~ expensive,data=HMO_data2)
```





