---
title: "Linear_Model"
author: "Shweta Rane"
date: "2022-11-27"
---

```{r}
library(tidyverse)
```

```{r}
#This file contains HMO data with outcome variable "expensive" derived using K-means
data <- read.csv("HMO_labeled_data.csv")
```

```{r}
data <- as.data.frame(data)
```

```{r}
str(data)
```
```{r}
data_new <- data.frame(id=as.numeric(data$X),
                     age=as.numeric(data$age),
                     bmi=as.numeric(data$bmi),
                     children=as.numeric(data$children),
                     smoker=as.numeric(ifelse(data$smoker == 'yes',1,0)),
                     location=as.numeric(unclass(as.factor(data$location))),
                     location_type=as.numeric(ifelse(data$location_type == 'Country',1,0)),
                     education_level=as.numeric(unclass(as.factor(data$education_level))),
                     yearly_physical=as.numeric(ifelse(data$yearly_physical == 'Yes',1,0)),
                     exercise=as.numeric(ifelse(data$exercise == 'Active',1,0)),
                     married=as.numeric(ifelse(data$married == 'Married',1,0)),
                     hypertension = as.numeric(data$hypertension),
                     gender=as.numeric(ifelse(data$gender == 'male',1,0)),
                     cost=as.numeric(data$cost))
                     
                     
```

```{r}
head(data_new)
```
```{r}
data_new[c('age','bmi','children','location','education_level','cost')] <- 
  scale(data_new[c('age','bmi','children','location','education_level','cost')])
```


```{r}
head(data_new)
```
```{r}
data_new2 <- data_new[c('age','bmi', 'smoker', 'cost' , 'children', 'id', 'location', 'location_type', 'yearly_physical', 'education_level', 'exercise', 'married', 'hypertension', 'gender')]
head(data_new2)
```

```{r}
#Applied linear regression on outcome variable cost using predictor bmi
lmOut1 <- lm(formula = cost ~ bmi, data = data_new2)
summary(lmOut1)
```
```{r}
#Applied linear regression on outcome variable cost using predictor smoker
lmOut2 <- lm(formula = cost ~ smoker, data = data_new2)
summary(lmOut2)
```
```{r}
#Applied linear regression on outcome variable cost using predictor age
lmOut3 <- lm(formula = cost ~ age, data = data_new2)
summary(lmOut3)
```

```{r}
#Applied linear regression on outcome variable cost using predictors bmi & age
lmOut4 <- lm(formula = cost ~ bmi + age, data = data_new2)
summary(lmOut4)

```
```{r}
#Applied linear regression on outcome variable cost using predictors bmi & smoker
lmOut5 <- lm(formula = cost ~ bmi + smoker, data = data_new2)
summary(lmOut5)
```
```{r}
#Applied linear regression on outcome variable cost using predictors smoker and age
lmOut6 <- lm(formula = cost ~ age + smoker, data = data_new2)
summary(lmOut6)
```
```{r}
#Applied linear regression on outcome variable cost using predictors bmi, smoker and age
lmOut7 <- lm(formula = cost ~ bmi + age + smoker, data = data_new2)
summary(lmOut7)
```
```{r}
#Applied linear regression on outcome variable cost using predictors bmi, smoker, age, children, location, location_type, education_level, exercise, married, hypertension, gender, yearly_physical, id
lmOut8 <- lm(formula = cost ~ bmi + age + smoker + children +location + location_type + exercise + married + hypertension + gender + education_level + yearly_physical , data = data_new2)
summary(lmOut8)
```
```{r}
#Applied linear regression on outcome variable cost using predictors bmi, smoker, age, children,exercise, hypertension 
lmOut9 <- lm(formula = cost ~ bmi + age + smoker + children + exercise  + hypertension , data = data_new2)
summary(lmOut9)
```

```{r}
#Creating a dataframe using the suggested values
predDF <- data.frame(age = 0.5723449, bmi=0.4422956, smoker=0)
predict(lmOut7, predDF)
```
We tried applying linear model using various parameters and combinations using  bmi, smoker, age, children, location, location_type, education_level, exercise, married, hypertension, gender, yearly_physical, id to predict cost. By adding the independent variable i.e bmi, age, smoker, children, exercise and hypertension we are increasing the overall prediction capacity of our model which is 57%.


