---
title: "Linear_Model"
author: "Shweta Rane"
date: "2022-11-27"
---

```{r}
#This file contains HMO data with outcome variable "expensive" derived using K-means
library(tidyverse)
file <- ("https://intro-datascience.s3.us-east-2.amazonaws.com/HMO_data.csv")
data <- read_csv(file)
head(data)
```

```{r}
data <- as.data.frame(data)
```

```{r}
#install.packages("imputeTS")
library(imputeTS)
data$bmi <- na_interpolation(data$bmi)
#View(HMO_data)
sum(is.na(data$bmi))

data$hypertension <- na_interpolation(data$hypertension)
#View(HMO_data)
sum(is.na(data$hypertension))
```

```{r}
str(data)
```
```{r}
data_new <- data.frame(id=as.numeric(data$X),
                     age=as.numeric(data$age),
                     bmi=as.numeric(data$bmi),
                     children=as.numeric(data$children),
                     smoker=as.numeric(ifelse(data$smoker == 'yes',1,0)),
                     location=as.numeric(unclass(as.factor(data$location))),
                     location_type=as.numeric(ifelse(data$location_type == 'Country',1,0)),
                     education_level=as.numeric(unclass(as.factor(data$education_level))),
                     yearly_physical=as.numeric(ifelse(data$yearly_physical == 'Yes',1,0)),
                     exercise=as.numeric(ifelse(data$exercise == 'Active',1,0)),
                     married=as.numeric(ifelse(data$married == 'Married',1,0)),
                     hypertension = as.numeric(data$hypertension),
                     gender=as.numeric(ifelse(data$gender == 'male',1,0)),
                     cost=as.numeric(data$cost))
                     
                     
```

```{r}
data_new
head(data_new)
```
```{r}
# install.packages('corrplot')
library(corrplot)
corrmatrix <- cor(data_new)

corrplot(cor(corrmatrix),addCoef.col="black",number.cex=0.30)
```
```{r}
plot_1 <- ggplot(data=data_new) + aes(y=cost, x=age) + geom_point() + #making a visualization
geom_smooth(method="lm", se=FALSE)  #giving a line of best fit to the scatter plot

plot_2 <- ggplot(data=data_new) + aes(y=cost, x=bmi) + geom_point() + #making a visualization
geom_smooth(method="lm", se=FALSE)


plot_3 <- ggplot(data=data_new) + aes(y=cost, x=children) + geom_point() + #making a visualization
geom_smooth(method="lm", se=FALSE)

plot_4 <- ggplot(data=data_new) + aes(y=cost, x=hypertension) + geom_point() + #making a visualization
geom_smooth(method="lm", se=FALSE)

plot_5 <- ggplot(data=data_new) + aes(y=cost, x=smoker) + geom_point() + #making a visualization
geom_smooth(method="lm", se=FALSE)


plot_1
plot_2
plot_3
plot_4
plot_5
```

```{r}
data_new[c('age','bmi','children','location','education_level','cost')] <- 
  scale(data_new[c('age','bmi','children','location','education_level','cost')])
```


```{r}
head(data_new)
```
```{r}
data_new2 <- data_new[c('age','bmi', 'smoker', 'cost' , 'children', 'id', 'location', 'location_type', 'yearly_physical', 'education_level', 'exercise', 'married', 'hypertension', 'gender')]
head(data_new2)
```

```{r}
#Applied linear regression on outcome variable cost using predictor bmi
lmOut1 <- lm(formula = cost ~ bmi, data = data_new2)
summary(lmOut1)
```
```{r}
#Applied linear regression on outcome variable cost using predictor smoker
lmOut2 <- lm(formula = cost ~ smoker, data = data_new2)
summary(lmOut2)
```
```{r}
#Applied linear regression on outcome variable cost using predictor age
lmOut3 <- lm(formula = cost ~ age, data = data_new2)
summary(lmOut3)
```

```{r}
#Applied linear regression on outcome variable cost using predictors bmi & age
lmOut4 <- lm(formula = cost ~ bmi + age, data = data_new2)
summary(lmOut4)

```
```{r}
#Applied linear regression on outcome variable cost using predictors bmi & smoker
lmOut5 <- lm(formula = cost ~ bmi + smoker, data = data_new2)
summary(lmOut5)
```
```{r}
#Applied linear regression on outcome variable cost using predictors smoker and age
lmOut6 <- lm(formula = cost ~ age + smoker, data = data_new2)
summary(lmOut6)
```
```{r}
#Applied linear regression on outcome variable cost using predictors bmi, smoker and age
lmOut7 <- lm(formula = cost ~ bmi + age + smoker, data = data_new2)
summary(lmOut7)
```
```{r}
#Applied linear regression on outcome variable cost using predictors bmi, smoker, age, children, location, location_type, education_level, exercise, married, hypertension, gender, yearly_physical, id
lmOut8 <- lm(formula = cost ~ bmi + age + smoker + children +location + location_type + exercise + married + hypertension + gender + education_level + yearly_physical , data = data_new2)
summary(lmOut8)
```
```{r}
#Applied linear regression on outcome variable cost using predictors bmi, smoker, age, children,exercise, hypertension 
lmOut9 <- lm(formula = cost ~ bmi + age + smoker + children + exercise  + hypertension , data = data_new2)
summary(lmOut9)
```

```{r}
#Creating a dataframe using the suggested values
predDF <- data.frame(age = 0.5723449, bmi=0.4422956, smoker=0)
predict(lmOut7, predDF)
```
We tried applying linear model using various parameters and combinations using  bmi, smoker, age, children, location, location_type, education_level, exercise, married, hypertension, gender, yearly_physical, id to predict cost. By adding the independent variable i.e bmi, age, smoker, children, exercise and hypertension we are increasing the overall prediction capacity of our model which is 57%.
```{r}
#upload the test file in the testdata below

testdata<- read.csv('C:\\Users\\Aruneema\\Desktop\\MSBA\\IST 687\\Project\\HMO_TEST_data_sample.csv')
#Creating a dataframe using the suggested values
#predDF <- data.frame(testdata)

predDF <- data.frame(id=as.numeric(testdata$X),
                     age=as.numeric(testdata$age),
                     bmi=as.numeric(testdata$bmi),
                     children=as.numeric(testdata$children),
                     smoker=as.numeric(ifelse(testdata$smoker == 'yes',1,0)),
                     location=as.numeric(unclass(as.factor(testdata$location))),
                     location_type=as.numeric(ifelse(testdata$location_type == 'Country',1,0)),
                     education_level=as.numeric(unclass(as.factor(testdata$education_level))),
                     yearly_physical=as.numeric(ifelse(testdata$yearly_physical == 'Yes',1,0)),
                     exercise=as.numeric(ifelse(testdata$exercise == 'Active',1,0)),
                     married=as.numeric(ifelse(testdata$married == 'Married',1,0)),
                     hypertension = as.numeric(testdata$hypertension),
                     gender=as.numeric(ifelse(testdata$gender == 'male',1,0))
                     )

predDF 

```

```{r}
#prediction
predict(lmOut9, predDF)

```

```{r}

```




